---
output:
  md_document:
    variant: markdown_github
---

# WHO

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/WHO)](http://cran.r-project.org/web/packages/WHO)
[![Travis-CI Build Status](https://travis-ci.org/expersso/WHO.svg?branch=master)](https://travis-ci.org/expersso/WHO)
[![codecov.io](https://codecov.io/github/expersso/WHO/coverage.svg?branch=master)](https://codecov.io/github/expersso/WHO?branch=master)
[![Cranlogs Downloads](http://cranlogs.r-pkg.org/badges/grand-total/WHO)](http://cran.r-project.org/web/packages/WHO)

```{r options, echo=FALSE}
knitr::opts_chunk$set(cache = FALSE, warning = FALSE, error = FALSE, 
                      fig.path = "")
library(WHO)
```

### Introduction

The `WHO` package allows the user to download public health data from the 
[World Health Organization's](http://www.WHO.int/) 
[Global Health Observatory](http://www.WHO.int/gho/en/) 
in a dynamic and reproducible way.

The package can be installed from either CRAN or Github (development version):

```{r install, eval=FALSE}
# From CRAN
install.packages("WHO")

# From Github
library(devtools)
install_github("expersso/WHO")

library(WHO)
```

### Usage Example

The `get_codes` function returns a data frame with series codes and descriptions
for all available series:

```{r get_codes}
library(dplyr)

codes <- get_codes()
head(codes)
```

(To retrieve additional meta information (e.g. French and Spanish descriptions,
category breakdowns of series, etc), use `get_codes(extra = TRUE)`.)

To find a series of interest, use either `View(codes)` in Rstudio, or search
with regular expressions:

```{r find_series}
codes[grepl("[Ll]ife expectancy", codes$display), ]
```

The `codes` data frame also provides a `url` to the meta data for a specified
series:

```{r, eval=FALSE}
# Opens a browser with the meta data for the specified series
browseURL(codes$url[1])
```

Having found the series of interest (in the `label` column), we can easily
retrieve the data and, for example, make a chart:

```{r example_1, fig.width=5, fig.height=3}
library(ggplot2)

df <- get_data("WHOSIS_000001")

head(df)

df %>% 
  filter(sex == "Both sexes") %>% 
  group_by(region, year) %>%
  summarise(value = mean(value)) %>% 
  ggplot(aes(x = year, y = value, color = region, linetype = region)) +
  geom_line(size = 1) +
  theme_light(9) +
  labs(x = NULL, y = "Life expectancy at birth (years)\n", 
       linetype = NULL, color = NULL,
       title = "Evolution of life expectancy (by region)\n")
```


### Disclaimer

This package is in no way officially related to or endorsed by the WHO.
